### 2022-05-27 main/chapter_2ljets.tex
fill in ptmiss; add sec prefixes

\subsubsection{$b$-tagging}
\label{sec:btagging}
Tagging adds labels to objects.
Unlike how electrons and muons are very cleanly separated however, jet tagging
is a more noisy affair.
Taggers are available for various jet properties, such charm flavor ($c$-jet),
bottom flavor ($b$-jet).
Of these, $\twoljets$ uses only $b$-tagging.

@@

\subsection{Overlap removal}
Leptons and jets are independently reconstructed.
This means that, despite the various quality and isolation requirement, there
can remain examples where detector responses can be attributed to more than
one object.
For example, a track could be attributed to both a muon and an electron,
or a calorimeter energy cluster could be attributed to both an electron and a
photon. (Although we do not directly use photons in $\twoljets$, we do not veto
events containing them --- photons contribute to $\ptmiss$.)

@@

To remove duplicates and further reduce non-prompt lepton contributions, a
standard overlap removal procedure is applied.
This procedure, which is detailed in the paper~\cite{atlas2022searches},
comprises a sequence of rules which modify the collection of event objects.
Some rules decide which to keep from an overlapping pair,
and others remove leptons overlapping with jets, based on their $\Delta R$
and $\pt$.

\subsection{$\ptmiss$}

\subsubsection{$\met$ significance}
\label{sec:metsig}

Most common Standard Model processes at the LHC do not produce hard
invisible particles.
Our supersymmetric signals, however, do.
This makes the missing transverse momentum $\ptmiss$ its magnitude $\met$
important features for separating signal and background events in all parts
of the $\twoljets$ search.

@@

Uncertainty in $\ptmiss$ is modelled for $\metsig$ by a covariance matrix
$\Omega$ in the transverse plane, and used to define
\begin{equation}
\metsig
=
\ptmiss \Omega^{-1} \ptmiss
=
\frac{(\met)^2}{\sigma_L^2(1 - \rho_{LT}^2)}
,
\end{equation}
in which $\sigma_L^2$ is the variance in the direction of $\ptmiss$ and
$\rho_{LT}$ is the correlation coefficient between the parallel and
perpendicular components~\cite{atlas_met_significance}.
This is the square of a Mahalanobis (standard-deviations) distance between
the origin and $\ptmiss$~\cite{mahalanobis1936generalised}.

@@

\subsection{Stranverse mass $\mttwo$}
\label{sec:mt2}
Just as the slepton is a supersymmetric partner to a lepton, the stransverse
mass  $\mttwo$ is a partner to the `transverse mass' $m_T$
(to be defined shortly),
that is useful in searches for supersymmetric particles.

@@

\subsection{Cheat-sheet}
Uncomfortably many event variables are used to define the regions of this
analysis.
To aid future revision of their meanings, definitions and references for these
variables are collected here.
\begin{itemize}
\item $\met$: missing transverse momentum
\item $\metsig$: object-based missing transverse momentum significance;\\
described in Section~\ref{sec:metsig}
\item $\ptjone$: transverse momentum of the hardest jet%
\vspace{0.5em}
\item $\mll$: invariant mass of the hardest two leptons
\item $\mjj$: invariant mass of the two hardest jets
\item $\mbb$: invariant mass of the two hardest b-tagged jets
\item $\mjetone$: mass of the hardest jet
\item $\mttwoll$: `stransverse` mass of the lepton pair with massless invisibles;\\
described in Section~\ref{sec:mt2}
\vspace{0.5em}
\item $\rll$: distance in $\eta\textrm{--}\phi$ between the lepton pair
\item $\rjj$: distance in $\eta\textrm{--}\phi$ between the two hardest jets%
\vspace{0.5em}
\item $\njet$: number of signal jets
\item $\nbtag$: number of b-tagged jets;
described in Section~\ref{sec:btagging}
\vspace{0.5em}
\item $\dphillmet$: azimuthal angle between the lepton pair and $\ptmiss$
\item $\dphijmet$: azimuthal angle between the hardest jet and $\ptmiss$
\end{itemize}
All of these are physical properties of particles or collections of particles.
Please remember that none is an exact or true value;
all are are estimates based on approximate reconstructions.
While this distinction is easy to drop in lax language, is necessary to
understand the shapes of noisy variables such as $\mjj$ and indeed the
existence of $\metsig$.

>>>>

\subsubsection{$b$-tagging}
\label{sec:2ljets_btagging}
Tagging adds labels to objects.
Unlike how electrons and muons are very cleanly separated however, jet tagging
is a more noisy affair.
Taggers are available for various jet properties, such charm flavor ($c$-jet),
bottom flavor ($b$-jet).
Of these, $\twoljets$ uses only $b$-tagging.

@@

\subsection{Overlap removal}
\label{sec:2ljets_overlap_removal}
Leptons and jets are independently reconstructed.
This means that, despite the various quality and isolation requirement, there
can remain examples where detector responses can be attributed to more than
one object.
For example, a track could be attributed to both a muon and an electron,
or a calorimeter energy cluster could be attributed to both an electron and a
photon. (Although we do not directly use photons in $\twoljets$, we do not veto
events containing them --- photons contribute to $\ptmiss$.)

@@

To remove duplicates and further reduce non-prompt lepton contributions, we
apply a standard overlap removal procedure.
This procedure, which is detailed in the paper~\cite{atlas2022searches},
comprises a sequence of rules which modify the collection of event objects.
Some rules decide which to keep from an overlapping pair,
and others remove leptons overlapping with jets, based on their $\Delta R$
and $\pt$.

\subsection{$\ptmiss$}
Every event has $\ptmiss$.
Unlike leptons, jets, and other objects which we must select for, the $\ptmiss$
of an event is a natural result of the conservation of momentum, which \atlas'
barrel-shaped construction allows us to measure with reasonable precision.

The magnitude of $\ptmiss$ is $\met$, pronounced `met'.
Please note that $\met$ is not an energy.
Transverse energy does not exist.
One may of course define their own directed `transverse energy', but it will
not be conserved (in the standard model) unless it is equivalent to a conserved
momentum.
Indeed, momentum-balanced events do have energy imbalances;
for example, in a single-top event $pp \rightarrow tq$, the $t$ and $q$ are
produced back-to-back with equal momenta $p$, but the two particles have very
different energies since they have very different masses:
$E_t^2 = m_t^2 + p^2$, $E_q^2 = m_q^2 + p^2$, and $m_t \gg m_q$.
% hobbyhorse

The $\ptmiss$ vector is defined such that the vector sum of $\ptmiss$ with
all reconstructed objects (after overlap removal) has zero transverse momentum.
In addition to the leptons and jets used directly in $\twoljets$, this includes
photons, tau leptons, and a `soft term' comprising tracks not associated to
any other objects~\cite{atlas_met}.

Since signal models in $\twoljets$ do generate hard, invisible particles,
by selecting for large $\met$ we select events in which either there were real
invisible particles or something was badly mismeasured.
The standard selection for $\twoljets$-electroweak is $\met > 100\,\eV[G]$.
This is a nice round number, and corresponds roughly to where the $\met$-less
background $\zjets$ tails off.
This $100\,\eV[G]$ also corresponds roughly to several standard deviations of
a typical $\met$ noise scale of $\approx 10--30\,\eV[G]$, which I observe
by inspecting $\met / \metsig$ in $\twoljets$-electroweak selections;
$\metsig$ is defined in the next section.

\subsection{$\met$ significance}
\label{sec:2ljets_metsig}
Most common Standard Model processes at the LHC do not produce hard
invisible particles.
Our supersymmetric signals, however, do.
This makes the missing transverse momentum $\ptmiss$ its magnitude $\met$
important features for separating signal and background events in all parts
of the $\twoljets$ search.

@@

Uncertainty in $\ptmiss$ is modelled for $\metsig$ by a covariance matrix
$\Omega$ in the transverse plane, and used to define
\begin{equation}
\metsig^2
=
\ptmiss \Omega^{-1} \ptmiss
=
\frac{(\met)^2}{\sigma_L^2(1 - \rho_{LT}^2)}
,
\end{equation}
in which $\sigma_L^2$ is the variance in the direction of $\ptmiss$ and
$\rho_{LT}$ is the correlation coefficient between the parallel and
perpendicular components~\cite{atlas_met_significance}.
This is the square of a Mahalanobis (standard-deviations) distance between
the origin and $\ptmiss$~\cite{mahalanobis1936generalised}.

@@

\subsection{Stranverse mass $\mttwo$}
\label{sec:2ljets_mt2}
Just as the slepton is a supersymmetric partner to a lepton, the stransverse
mass  $\mttwo$ is a partner to the `transverse mass' $m_T$
(to be defined shortly),
that is useful in searches for supersymmetric particles.

@@

\subsection{Cheat-sheet}
Uncomfortably many event variables are used to define the regions of this
analysis.
To aid future revision of their meanings, definitions and references for these
variables are collected here.
\begin{itemize}
\item $\met$: missing transverse momentum
\item $\metsig$: object-based missing transverse momentum significance;\\
described in Section~\ref{sec:2ljets_metsig}
\item $\ptjone$: transverse momentum of the hardest jet%
\vspace{0.5em}
\item $\mll$: invariant mass of the hardest two leptons
\item $\mjj$: invariant mass of the two hardest jets
\item $\mbb$: invariant mass of the two hardest b-tagged jets
\item $\mjetone$: mass of the hardest jet
\item $\mttwoll$: `stransverse` mass of the lepton pair with massless invisibles;\\
described in Section~\ref{sec:2ljets_mt2}
\vspace{0.5em}
\item $\rll$: distance in $\eta\textrm{--}\phi$ between the lepton pair
\item $\rjj$: distance in $\eta\textrm{--}\phi$ between the two hardest jets%
\vspace{0.5em}
\item $\njet$: number of signal jets
\item $\nbtag$: number of b-tagged jets;
described in Section~\ref{sec:2ljets_btagging}
\vspace{0.5em}
\item $\dphillmet$: azimuthal angle between the lepton pair and $\ptmiss$
\item $\dphijmet$: azimuthal angle between the hardest jet and $\ptmiss$
\end{itemize}
All of these are physical properties of particles or collections of particles.
Please remember that none is an exact or true value;
all are are estimates based on approximate reconstructions.
While this distinction is easy to drop in lax language, is necessary to
understand the shapes of noisy variables such as $\mjj$ and indeed the
existence of $\metsig$.
